\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{enumitem}
\newtheorem{problem}{Problem}[section]

\title{MA2008B: Assignment 03}
\author{Dr. Juliho Castillo Colmeares}
\date{Period 03}

\begin{document}

\maketitle

\section*{Instructions}
\begin{itemize}
    \item Show all your work for full credit.
    \item Write your answers clearly and neatly.
    \item Submit your assignment by the due date.
\end{itemize}

\section{Partial Differential Equations}

\begin{problem}
Show that the solution to the initial value problem is unique 
provided that it is sufficiently smooth and decays sufficiently 
fast at infinity, as follows:

Suppose that $u_1(x, t)$ and $u_2(x, t)$ are both solutions to the 
initial value problem (4.5)--(4.9): 

\[\frac{\partial u}{\partial t} = \frac{\partial^2 u}{\partial x^2} - \infty < x < \infty, \quad t > 0, \tag{4.5}\]

with

\[u(x, 0) = u_0(x), \tag{4.6}\]

where

(i) \( u_0(x) \) is sufficiently well behaved, \hspace{20pt} (4.7)

(ii) \[ \lim_{|x|\to\infty} u_0(x)e^{-ax^2} = 0 \] for any \( a > 0 \), \hspace{20pt} (4.8)

and lastly where

\[ \lim_{|x|\to\infty} u(x,t)e^{-ax^2} = 0 \] for any \( a > 0 \), \( t > 0 \), \hspace{20pt} (4.9)

Show that $v(x, t) = u_1 - u_2$ 
is also a solution of (4.5) with $v(x, 0) = 0$. 

Show that if
\[
E(t) = \int_{-\infty}^\infty v^2 \, dx,
\]
then
\[
E(t) \geq 0, \quad E(0) = 0,
\]
and, by integrating by parts, that
\[
\frac{dE}{dt} \leq 0;
\]
thus $E(t) \equiv 0$, hence $v(x, t) \equiv 0$.

Note, though, that as yet we have no guarantee that $v(x, t)$ exists, 
nor that the above manipulations can be justified.
\end{problem}

\begin{problem}
Show that \(\sin nx \, e^{-n^2t}\) is a solution of the forward diffusion equation, and that \(\sin nx \, e^{n^2t}\) is a solution of the backward diffusion equation. Now try to solve the initial value problem for the forward and backward equations in the interval \(-\pi < x < \pi\) by expanding the solution in a Fourier series in \(x\) with coefficients depending on \(t\). What difference do you see between the two problems? Which is well-posed?

    
\end{problem}

\begin{thebibliography}{9}

\bibitem{wilmott}
P. Wilmott, J. Dewynne, S. Howison,
\textit{Option Pricing: Mathematical Models and Computation},
Oxford Financial Press, 1993.

\end{thebibliography}

\end{document}